# Tumor: Sub-clustering on T/NK lineage {#sec-tumor-tnk-subclustering}

## Set up Seurat workspace

```{r }
#| warning: false
#| error: false

# Load libraries
library(data.table)
library(devtools)
library(presto)
library(glmGamPoi)
library(sctransform)
library(Seurat) 
library(tidyverse)
library(miQC)
library(SeuratWrappers)
library(flexmix)
library(SingleCellExperiment)
library(SummarizedExperiment)
library(readxl)
library(fishpond)
library(Matrix)
library(speckle)
library(scater)
library(patchwork)
library(vctrs)
library(alevinQC)
library(harmony)
library(scDblFinder)
library(cellXY)

# Set global options for Seurat v5 objects
options(Seurat.object.assay.version = 'v5')
```

## Load previously saved clustered object

```{r}
merged.18279.tumor.singlets <- readRDS("Tumor_scRNA_Part3.rds")
```

## Set idents to preferred initial clustering resolution

```{r}
Idents(merged.18279.tumor.singlets) <- merged.18279.tumor.singlets$RNA_snn_res.1
merged.18279.tumor.singlets$seurat_clusters <- merged.18279.tumor.singlets$RNA_snn_res.1
DimPlot(merged.18279.tumor.singlets, reduction="umap.harmony", label=T)
```

## Run sub-clustering on T/NK lineage
T/NK cells are in clusters 1,2,3,9,12,16,18,22,26

### Cluster 1

```{r}
merged.18279.tumor.singlets <- FindSubCluster(object = merged.18279.tumor.singlets, 
										cluster = "1",
										resolution = 0.2,
										graph.name = "RNA_snn",  
										algorithm = 2)

DimPlot(merged.18279.tumor.singlets,  
        reduction = "umap.harmony",  
        group.by = "sub.cluster", 
        label = TRUE)

table(merged.18279.tumor.singlets$sub.cluster)
Idents(merged.18279.tumor.singlets) <- "sub.cluster"
levels(merged.18279.tumor.singlets)
```

### Cluster 2

```{r}
merged.18279.tumor.singlets <- FindSubCluster(object = merged.18279.tumor.singlets, 
										cluster = "2",
										resolution = 0.2,
										graph.name = "RNA_snn",  
										algorithm = 2)

DimPlot(merged.18279.tumor.singlets,  
        reduction = "umap.harmony",  
        group.by = "sub.cluster", 
        label = TRUE)

table(merged.18279.tumor.singlets$sub.cluster)
Idents(merged.18279.tumor.singlets) <- "sub.cluster"
levels(merged.18279.tumor.singlets)
```

### Cluster 3

```{r}
merged.18279.tumor.singlets <- FindSubCluster(object = merged.18279.tumor.singlets, 
										cluster = "3",
										resolution = 0.2,
										graph.name = "RNA_snn",  
										algorithm = 2)

DimPlot(merged.18279.tumor.singlets,  
        reduction = "umap.harmony",  
        group.by = "sub.cluster", 
        label = TRUE)

table(merged.18279.tumor.singlets$sub.cluster)
Idents(merged.18279.tumor.singlets) <- "sub.cluster"
levels(merged.18279.tumor.singlets)
```

### Cluster 9

```{r}
merged.18279.tumor.singlets <- FindSubCluster(object = merged.18279.tumor.singlets, 
										cluster = "9",
										resolution = 0.2,
										graph.name = "RNA_snn",  
										algorithm = 2)

DimPlot(merged.18279.tumor.singlets,  
        reduction = "umap.harmony",  
        group.by = "sub.cluster", 
        label = TRUE)

table(merged.18279.tumor.singlets$sub.cluster)
Idents(merged.18279.tumor.singlets) <- "sub.cluster"
levels(merged.18279.tumor.singlets)
```

### Cluster 12

```{r}
merged.18279.tumor.singlets <- FindSubCluster(object = merged.18279.tumor.singlets, 
										cluster = "12",
										resolution = 0.2,
										graph.name = "RNA_snn",  
										algorithm = 2)

DimPlot(merged.18279.tumor.singlets,  
        reduction = "umap.harmony",  
        group.by = "sub.cluster", 
        label = TRUE)

table(merged.18279.tumor.singlets$sub.cluster)
Idents(merged.18279.tumor.singlets) <- "sub.cluster"
levels(merged.18279.tumor.singlets)
```

### Cluster 16

```{r}
merged.18279.tumor.singlets <- FindSubCluster(object = merged.18279.tumor.singlets, 
										cluster = "16",
										resolution = 0.2,
										graph.name = "RNA_snn",  
										algorithm = 2)

DimPlot(merged.18279.tumor.singlets,  
        reduction = "umap.harmony",  
        group.by = "sub.cluster", 
        label = TRUE)

table(merged.18279.tumor.singlets$sub.cluster)
Idents(merged.18279.tumor.singlets) <- "sub.cluster"
levels(merged.18279.tumor.singlets)
```

### Cluster 18

```{r}
merged.18279.tumor.singlets <- FindSubCluster(object = merged.18279.tumor.singlets,
                                                                                cluster = "18",
                                                                                resolution = 0.2,
                                                                                graph.name = "RNA_snn",
                                                                                algorithm = 2)

DimPlot(merged.18279.tumor.singlets,
        reduction = "umap.harmony",
        group.by = "sub.cluster",
        label = TRUE)

table(merged.18279.tumor.singlets$sub.cluster)
Idents(merged.18279.tumor.singlets) <- "sub.cluster"
levels(merged.18279.tumor.singlets)
```

### Cluster 22

```{r}
merged.18279.tumor.singlets <- FindSubCluster(object = merged.18279.tumor.singlets,
                                                                                cluster = "22",
                                                                                resolution = 0.2,
                                                                                graph.name = "RNA_snn",
                                                                                algorithm = 2)

DimPlot(merged.18279.tumor.singlets,
        reduction = "umap.harmony",
        group.by = "sub.cluster",
        label = TRUE)

table(merged.18279.tumor.singlets$sub.cluster)
Idents(merged.18279.tumor.singlets) <- "sub.cluster"
levels(merged.18279.tumor.singlets)
```

### Cluster 26

```{r}
merged.18279.tumor.singlets <- FindSubCluster(object = merged.18279.tumor.singlets,
                                                                                cluster = "26",
                                                                                resolution = 0.2,
                                                                                graph.name = "RNA_snn",
                                                                                algorithm = 2)

DimPlot(merged.18279.tumor.singlets,
        reduction = "umap.harmony",
        group.by = "sub.cluster",
        label = TRUE)

table(merged.18279.tumor.singlets$sub.cluster)
Idents(merged.18279.tumor.singlets) <- "sub.cluster"
levels(merged.18279.tumor.singlets)
```

## Plot updated QC metrics per cluster/sub-cluster

```{r}
#| fig-width: 15
VlnPlot(merged.18279.tumor.singlets, features = "nCount_RNA", group.by="sub.cluster", pt.size=0) + NoLegend()
VlnPlot(merged.18279.tumor.singlets, features = "nFeature_RNA", group.by="sub.cluster", pt.size=0) + NoLegend()
VlnPlot(merged.18279.tumor.singlets, features = "percent.mt", group.by="sub.cluster", pt.size=0) + NoLegend()
```

## Identify updated cursory marker genes per cluster/sub-cluster

```{r}
DefaultAssay(merged.18279.tumor.singlets) <- "RNA"

vargenes <- presto::wilcoxauc(merged.18279.tumor.singlets, 'sub.cluster', seurat_assay = 'RNA')
top_vargenes <- top_markers(vargenes, n = 100, auc_min = 0.5, pct_in_min = 50, pct_out_max = 50)
top_vargenes

write_tsv(top_vargenes,"Tumor_scRNA_prestoMarkers.tsv")

```

## Plot top markers identified and canonical genes as a dotplot

```{r}
#| fig-width: 35
#| fig-height: 10
tumor <- c("DCT","MLANA","MITF","PMEL","S100A1","TYR","APOC1")
endothelial <- c("PECAM1","VWF")
fibroblast <- c("COL3A1","COL1A1","COL1A2","LUM")
tcell <- c("FGFBP2","KLRD1","CD3E","CD3D","GZMB","XCL2","GZMH","CST7","GZMK","GZMA","IFNG","GNLY","CCL4","NKG7","CCL5","CD8A","CD8B","CTLA4","TNFRSF4","BATF","ITM2A")
mono <- c("LYZ","CD74","CD68")
bcell <- c("MS4A1","CD79A")

top_vargenes <- top_markers(vargenes, n = 5, auc_min = 0.5, pct_in_min = 50, pct_out_max = 50)
top_markers <- top_vargenes %>%
    select(-rank) %>% 
    unclass() %>% 
    stack() %>%
    pull(values) %>%
    unique() %>%
    .[!is.na(.)]

dotplotmarkers <- unique(c(top_markers,tumor,endothelial,fibroblast,tcell,mono,bcell))

# Compute aggregated expression values of these genes and cluster them to order the figure
rna <- AverageExpression(merged.18279.tumor.singlets,assay="RNA",slot="data")
rna.sub <- rna$RNA[dotplotmarkers,]
cors.genes <- as.dist(1-cor(as.matrix(t(rna.sub)),method="pearson"))
hc.genes <- hclust(cors.genes)
dotplotmarkers.sorted <- rownames(rna.sub)[hc.genes$order]

# Also re-order the sub-clusters
cors.subs <- as.dist(1-cor(as.matrix(rna.sub),method="pearson"))
hc.subs <- hclust(cors.subs)
merged.18279.tumor.singlets@active.ident <- factor(merged.18279.tumor.singlets@active.ident, 
                            levels=str_replace_all(str_replace_all(colnames(rna.sub)[hc.subs$order],"g",""),"-","_"))

# Plot
DotPlot(merged.18279.tumor.singlets,
	features = dotplotmarkers.sorted,
	assay = "RNA",
	cols = c("blue","red")
	) + 
	RotatedAxis()

```

## Plot updated tumor proportions per cluster over time

```{r}
#| fig-width: 10
#| fig-height: 8
# Tumor samples, week 0, 12, 20 timepoints
as.data.frame(table(merged.18279.tumor.singlets$sub.cluster, merged.18279.tumor.singlets$Sample)) %>% 
  as_tibble() %>%
  dplyr::rename("Sample" = Var2,"Cluster" = Var1) %>%
  separate(Sample, into=c("Patient","Site","Timepoint","IpiCohort","Assay"),sep="_",remove=F) %>%
  dplyr::filter(Timepoint=="W00" | Timepoint=="W12" | Timepoint=="W20") %>%
  group_by(Sample) %>%
  mutate(Proportion = Freq / sum(Freq)) %>%
  ggplot(aes(fill = Timepoint, x = Timepoint, y=Proportion)) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width=0.2)) +
  facet_wrap(~Cluster,scales="free")
```

## Save updated object

```{r}
saveRDS(merged.18279.tumor.singlets, "Tumor_scRNA_Part6.rds")
```


## Get session info

```{r}
sessionInfo()
```
