# Save scTCR and bulkTCR colorschemes {#sec-pbmc-define-colorschemes}

## Set up workspace
```{r}
# Libraries
library(Seurat)
library(colorblindr)
library(RColorBrewer)
library(paletteer)
library(dplyr)
library(tidyverse)
```

## Assign custom color scheme
```{r}
projectils_fill <- scale_fill_manual(values = c("CD4.NaiveLike" = "#F7FBFF", "CD4.Memory" = "#DEEBF7", "CD4.Tfh" = "#C6DBEF", "CD4.Th17" = "#9ECAE1", "CD4.CTL_GNLY" = "#6BAED6", "CD4.CTL_EOMES" = "#4292C6", "CD4.CTL_Exh" = "#2171B5", "CD4.Treg" = "#084594", "CD8.NaiveLike" = "#FEE5D9", "CD8.MAIT" = "#FCBBA1", "CD8.CM" = "#FC9272", "CD8.EM" = "#FB6A4A", "CD8.TEMRA" = "#EF3B2C", "CD8.TPEX" = "#CB181D", "CD8.TEX" = "#99000D"))

projectils_color <- scale_color_manual(values = c("CD4.NaiveLike" = "#F7FBFF", "CD4.Memory" = "#DEEBF7", "CD4.Tfh" = "#C6DBEF", "CD4.Th17" = "#9ECAE1", "CD4.CTL_GNLY" = "#6BAED6", "CD4.CTL_EOMES" = "#4292C6", "CD4.CTL_Exh" = "#2171B5", "CD4.Treg" = "#084594", "CD8.NaiveLike" = "#FEE5D9", "CD8.MAIT" = "#FCBBA1", "CD8.CM" = "#FC9272", "CD8.EM" = "#FB6A4A", "CD8.TEMRA" = "#EF3B2C", "CD8.TPEX" = "#CB181D", "CD8.TEX" = "#99000D"))
```

## Relevel functional.cluster factor before plotting
```{r}
projectils_celltypes <- c("CD4.NaiveLike", "CD4.Memory", "CD4.Tfh", "CD4.Th17", "CD4.CTL_GNLY", "CD4.CTL_EOMES", "CD4.CTL_Exh", "CD4.Treg", "CD8.NaiveLike","CD8.MAIT", "CD8.CM", "CD8.EM", "CD8.TEMRA", "CD8.TPEX", "CD8.TEX")

projectils_celltypes <- factor(projectils_celltypes, levels = c("CD4.NaiveLike", "CD4.Memory", "CD4.Tfh", "CD4.Th17", "CD4.CTL_GNLY", "CD4.CTL_EOMES", "CD4.CTL_Exh", "CD4.Treg", "CD8.NaiveLike","CD8.MAIT", "CD8.CM", "CD8.EM", "CD8.TEMRA", "CD8.TPEX", "CD8.TEX"))
```

## Check plots are working
```{r}
# A plot with all the data
p <- projectils_celltypes %>%
  as.data.frame(functional.cluster = projectils_celltypes) %>%
  ggplot(aes(x = 1, fill = projectils_celltypes)) + 
  geom_bar() + 
  projectils_fill
p

```

## Check colors for colorblindness
```{r}
cvd_grid(p)
```

## Save color and fill environment variables
```{r}
saveRDS(projectils_fill, "Part0_ProjecTILs_Fill.rds")
saveRDS(projectils_color, "Part0_ProjecTILs_Color.rds")
```

## Load ifng, tumor and skin single cell data
```{r}
patients <- c("P101", "P103", "P104", "P105", "P106", "P108", "P109", "P110", "P111")
```

## Create factors
```{r}
patients <- data.frame(Patient = patients) %>%
  mutate(Patient = factor(Patient))
```

## Assign custom color scheme
```{r}
colors <- c( "#59A14F",  "#B07AA1FF", "#76B7B2FF", "#FBB258FF", "#DC050C", "#F6AAC9FF", "#5ca2e5", "#615ebf", "#826250") 

names(colors) <- levels(patients$Patient)
patient_fill <- scale_fill_manual(values = colors)
patient_color <- scale_color_manual(values = colors)
```

## Check plots are working
```{r}
# A plot with all the data
p <- ggplot(patients, aes(x = 1,fill = Patient)) + 
  geom_bar() + 
  patient_fill
p + ggtitle("BulkTCR Patients")

# A plot with the single cell patients: P101, P103, P104 and P108
p2 <- patients[c(1,2,3,6), 1, drop = FALSE] %>%
  ggplot(aes(x = 1,fill = Patient)) + 
  geom_bar() + 
  patient_fill
p2 + ggtitle("scRNA Patients")
```

## Check colors for colorblindness
```{r}
cvd_grid(p)
cvd_grid(p2)
```

## Save color and fill environment variables
```{r}
saveRDS(patient_fill, "Part0_Patient_Fill.rds")
saveRDS(patient_color, "Part0_Patient_Color.rds")
```

## Get session info
```{r}
sessionInfo()
```

