# scTCRseq: Plot IMP expanded and scTIL beta chain overlap {sec-sctcr-imp-expanded-venn-diagram}

## Set up workspace
```{r}
# Libraries
library(tidyverse)
library(Seurat)
library(dplyr)
library(scRepertoire)
library(ggvenn)
```

## Open mapping of single cell clonotypes to separated alpha and beta chains from scTCR data (regardless of scRNA)
```{r}
sc_bulk_mapping <- read.csv("scTCR_bulkTCR_chain_mapping_Part4.csv")
```

## Load categories
```{r}
p101_til_categories <- read.csv("sctcr_scRep_p101_til_freq_Part2.csv") %>%
  mutate(Patient = "P101")
p103_til_categories <- read.csv("sctcr_scRep_p103_til_freq_Part2.csv") %>%
  mutate(Patient = "P103")

til_categories <- rbind(p101_til_categories[,c("Patient", "vjaa", "sctcr_category")], p103_til_categories[,c("Patient", "vjaa", "sctcr_category")])
```

## Join categories
```{r}
sc_bulk_mapping <- sc_bulk_mapping %>%
  left_join(til_categories) %>%
  filter(tumor_vjaa_counts > 0)

sc_bulk_mapping %>%
  group_by(Patient, sctcr_category) %>%
  dplyr::count()
```

## Load IMP expanded clones
```{r}
p101_betas_expanded <- read.csv("p101_betas_merged_typed_imp_expanded_Part3.csv") %>%
  mutate(Beta_clonotype_Pt = paste0(Patient, "_", Beta_clonotype)) %>%
  dplyr::rename("bulk_Type" = "Type")
p103_betas_expanded <- read.csv("p103_betas_merged_typed_imp_expanded_Part3.csv") %>%
  mutate(Beta_clonotype_Pt = paste0(Patient, "_", Beta_clonotype)) %>%
  dplyr::rename("bulk_Type" = "Type")
```

## Count all IMP expanded beta chains in bulk
```{r}
imp_expanded_betas_bulk <- rbind(p101_betas_expanded[,c("Patient", "Beta_clonotype_Pt", "bulk_Type")], p103_betas_expanded[,c("Patient", "Beta_clonotype_Pt", "bulk_Type")])

imp_expanded_betas_bulk_count <- imp_expanded_betas_bulk %>%
  dplyr::count(Patient) %>%
  dplyr::rename("Total_IMPExpanded_Betas_Bulk" = "n")

```

## Count unique beta chains in single cell TILs per timepoint
```{r}
sctils_betas <- sc_bulk_mapping %>%
  filter(str_detect(bulk_chain, "TRB"),
         Patient %in% c("P101", "P103", "P104", "P108"),
         tumor_vjaa_counts > 0) %>%
  mutate(Beta_clonotype_Pt = paste0(Patient, "_", bulk_chain))

w00_sctils_betas <- sctils_betas %>%
  filter(sctcr_category == "Existing")

w12_sctils_betas <- sctils_betas %>%
  filter(sctcr_category == "Post-Nivolumab")

w20_sctils_betas <- sctils_betas %>%
  filter(sctcr_category == "Post-Vaccine")

sctils_betas_count <- sctils_betas %>%
  dplyr::count(Patient) %>%
  dplyr::rename("Total_Betas_scTILs" = "n")

w00_sctils_betas_count <- w00_sctils_betas %>%
  distinct(Patient, Beta_clonotype_Pt) %>%
  dplyr::count(Patient) %>%
  dplyr::rename("W00_Total_Betas_scTILs" = "n")

w12_sctils_betas_count <- w12_sctils_betas %>%
  distinct(Patient, Beta_clonotype_Pt) %>%
  dplyr::count(Patient) %>%
  dplyr::rename("W12_Total_Betas_scTILs" = "n")

w20_sctils_betas_count <- w20_sctils_betas %>%
  distinct(Patient, Beta_clonotype_Pt) %>%
  dplyr::count(Patient) %>%
  dplyr::rename("W20_Total_Betas_scTILs" = "n")
```

## Count IMP-expanded beta chains in common with scTIL beta chains per timepoint
```{r}
imp_expanded_common_betas <- inner_join(imp_expanded_betas_bulk, sctils_betas, by = c("Patient", "Beta_clonotype_Pt"))

imp_expanded_common_betas_count <- imp_expanded_common_betas %>%
  dplyr::count(Patient) %>%
  dplyr::rename("Common betas" = "n")

w00_common_betas <- inner_join(imp_expanded_betas_bulk, w00_sctils_betas, by = c("Patient", "Beta_clonotype_Pt"))

w00_common_betas_count <- w00_common_betas %>%
  distinct(Patient, Beta_clonotype_Pt) %>%
  dplyr::count(Patient) %>%
  dplyr::rename("Common betas" = "n")

w00_common_betas <- inner_join(imp_expanded_betas_bulk, w00_sctils_betas, by = c("Patient", "Beta_clonotype_Pt"))

w00_common_betas_count <- w00_common_betas %>%
  distinct(Patient, Beta_clonotype_Pt) %>%
  dplyr::count(Patient) %>%
  dplyr::rename("Common betas" = "n")

w12_common_betas <- inner_join(imp_expanded_betas_bulk, w12_sctils_betas, by = c("Patient", "Beta_clonotype_Pt"))

w12_common_betas_count <- w12_common_betas %>%
  distinct(Patient, Beta_clonotype_Pt) %>%
  dplyr::count(Patient) %>%
  dplyr::rename("Common betas" = "n")

w20_common_betas <- inner_join(imp_expanded_betas_bulk, w20_sctils_betas, by = c("Patient", "Beta_clonotype_Pt"))

w20_common_betas_count <- w20_common_betas %>%
  distinct(Patient, Beta_clonotype_Pt) %>%
  dplyr::count(Patient) %>%
  dplyr::rename("Common betas" = "n")
```

## Plot a venn diagram of IMP-expanded beta chains and scTIL beta chains per timepoint for Fig S8A
```{r, fig.width = 6, fig.height = 3}
# W00
w00_betas_overlap_df <- imp_expanded_betas_bulk_count %>%
  full_join(w00_sctils_betas_count) %>%
  full_join(w00_common_betas_count) %>%
  replace(is.na(.), 0) %>%
  rowwise() %>%
  mutate(`IMP expanded PBMC betas` = Total_IMPExpanded_Betas_Bulk-`Common betas`,
         `W00 scTIL betas` = W00_Total_Betas_scTILs-`Common betas`) %>%
  ungroup() 

# W12
w12_betas_overlap_df <- imp_expanded_betas_bulk_count %>%
  full_join(w12_sctils_betas_count) %>%
  full_join(w12_common_betas_count) %>%
  rowwise() %>%
  mutate(`IMP expanded PBMC betas` = Total_IMPExpanded_Betas_Bulk-`Common betas`,
         `W12 scTIL betas` = W12_Total_Betas_scTILs-`Common betas`) %>%
  ungroup() 

# W20
w20_betas_overlap_df <- imp_expanded_betas_bulk_count %>%
  full_join(w20_sctils_betas_count) %>%
  full_join(w20_common_betas_count) %>%
  rowwise() %>%
  mutate(`IMP expanded PBMC betas` = Total_IMPExpanded_Betas_Bulk-`Common betas`,
         `W20 scTIL betas` = W20_Total_Betas_scTILs-`Common betas`) %>%
  ungroup() 


w00_betas_overlap_formatted <- w00_betas_overlap_df %>%
  select(`IMP expanded PBMC betas`, `W00 scTIL betas`, `Common betas`, Patient) %>%
  pivot_longer(cols = c("IMP expanded PBMC betas", "W00 scTIL betas", "Common betas"), names_to = "Type", values_to = "n") %>%
  group_by(Patient) %>%
  mutate(Pt_sum = sum(n),
         percent = round(n/Pt_sum*100, 1),
         pct_label = case_when(Type == "IMP expanded PBMC betas" ~ paste0(percent, "%\n (", n, " clones)"),
                               Type == "W00 scTIL betas" ~ paste0(percent, "% (", n, " clones)"),
                               Type == "Common betas" ~ paste0(percent, "% (", n, " clones)"))) %>%
  ungroup() %>%
  mutate(Type = factor(Type, levels = c("IMP expanded PBMC betas", "Common betas", "W00 scTIL betas")),
         Patient = factor(Patient, levels = c("P108", "P104", "P103", "P101")))

w12_betas_overlap_formatted <- w12_betas_overlap_df %>%
  select(`IMP expanded PBMC betas`, `W12 scTIL betas`, `Common betas`, Patient) %>%
  pivot_longer(cols = c("IMP expanded PBMC betas", "W12 scTIL betas", "Common betas"), names_to = "Type", values_to = "n") %>%
  group_by(Patient) %>%
  mutate(Pt_sum = sum(n),
         percent = round(n/Pt_sum*100, 1),
         pct_label = case_when(Type == "IMP expanded PBMC betas" ~ paste0(percent, "%\n (", n, " clones)"),
                               Type == "W12 scTIL betas" ~ paste0(percent, "% (", n, " clones)"),
                               Type == "Common betas" ~ paste0(percent, "% (", n, " clones)"))) %>%
  ungroup() %>%
  mutate(Type = factor(Type, levels = c("IMP expanded PBMC betas", "Common betas", "W12 scTIL betas")),
         Patient = factor(Patient, levels = c("P108", "P104", "P103", "P101")))

w20_betas_overlap_formatted <- w20_betas_overlap_df %>%
  select(`IMP expanded PBMC betas`, `W20 scTIL betas`, `Common betas`, Patient) %>%
  pivot_longer(cols = c("IMP expanded PBMC betas", "W20 scTIL betas", "Common betas"), names_to = "Type", values_to = "n") %>%
  group_by(Patient) %>%
  mutate(Pt_sum = sum(n),
         percent = round(n/Pt_sum*100, 1),
         pct_label = case_when(Type == "IMP expanded PBMC betas" ~ paste0(percent, "%\n (", n, " clones)"),
                               Type == "W20 scTIL betas" ~ paste0(percent, "% (", n, " clones)"),
                               Type == "Common betas" ~ paste0(percent, "% (", n, " clones)"))) %>%
  ungroup() %>%
  mutate(Type = factor(Type, levels = c("IMP expanded PBMC betas", "Common betas", "W20 scTIL betas")),
         Patient = factor(Patient, levels = c("P108", "P104", "P103", "P101")))

w00_plot <- ggplot(w00_betas_overlap_formatted, aes(y = Patient, x = n, fill = Type)) +
  geom_col(position = "fill", width=0.5) +
  geom_text(aes(label = ifelse(Type == "Common betas", pct_label, "")), position = position_fill(vjust = 0.8), vjust = -2.5, size = 3) +
  geom_text(aes(label = ifelse(Type != "Common betas", pct_label, "")), position = position_fill(vjust = 0.5), size = 3) +
  theme_void() +
  theme(axis.text.y = element_text(size = 10, vjust = 1),
        plot.margin = margin(1, 1, 1, 1, "cm"),
        legend.position="bottom") +
  scale_fill_manual(values = c("grey", "darkblue", "beige")) +
  ggtitle("W00")

w12_plot <- ggplot(w12_betas_overlap_formatted, aes(y = Patient, x = n, fill = Type)) +
  geom_col(position = "fill", width=0.5) +
  geom_text(aes(label = ifelse(Type == "Common betas", pct_label, "")), position = position_fill(vjust = 0.8), vjust = -2.5, size = 3) +
  geom_text(aes(label = ifelse(Type != "Common betas", pct_label, "")), position = position_fill(vjust = 0.5), size = 3) +
  theme_void() +
  theme(axis.text.y = element_text(size = 10, vjust = 1),
        plot.margin = margin(1, 1, 1, 1, "cm"),
        legend.position="bottom") +
  scale_fill_manual(values = c("grey", "darkblue", "beige")) +
  ggtitle("W12")

w20_plot <- ggplot(w20_betas_overlap_formatted, aes(y = Patient, x = n, fill = Type)) +
  geom_col(position = "fill", width=0.5) +
  geom_text(aes(label = ifelse(Type == "Common betas", pct_label, "")), position = position_fill(vjust = 0.8), vjust = -2.5, size = 3) +
  geom_text(aes(label = ifelse(Type != "Common betas", pct_label, "")), position = position_fill(vjust = 0.5), size = 3) +
  theme_void() +
  theme(axis.text.y = element_text(size = 10, vjust = 1),
        plot.margin = margin(1, 1, 1, 1, "cm"),
        legend.position="bottom") +
  scale_fill_manual(values = c("grey", "darkblue", "beige")) +
  ggtitle("W20")

w00_plot
w12_plot
w20_plot

```

## Get session info
```{r}
sessionInfo()
```

