# scTCRseq: Mapping clones from scRNA and scTCRseq to bulkTCRseq {sec-map-sctcr-scrna-to-bulktcr}

## Set up workspace
```{r}
# Libraries
library(tidyverse)
library(Seurat)
library(dplyr)
library(scRepertoire)
```

## Load tumor data with clones loaded with a custom scRepertoire variable: vjaa (V gene, J gene, CDR3 aa)
```{r}
sc_tils <- readRDS("scRNA_TCR_Tumor_Part3.Rds")

sc_skin <- readRDS("scRNA_TCR_Skin_Part3.Rds")
```

## Create a dataframe of clones per patient, where each row is a chain from a clone.
Tumor
```{r}
# Splitting TRA and TRB
sctils_clone_chains <- sc_tils@meta.data %>%
  group_by(Patient) %>%
  dplyr::count(vjaa, Patient) %>%
  arrange(desc(n)) %>%
  dplyr::rename("tumor_vjaa_counts" = "n") %>%
  mutate(bulk_chain = strsplit(as.character(vjaa), "_")) %>%
  unnest(bulk_chain) %>%
  filter(bulk_chain != "NA;NA")

# Get clones with just 1 TRA or TRB chain
sctils_chain_one <- sctils_clone_chains %>%
  filter(str_count(bulk_chain, ";") == 1)

# Unnest clones with more than 1 TRA or TRB chain
sctils_chain_multi <- sctils_clone_chains %>%
  filter(str_count(bulk_chain, ";") > 1) %>%
  # split vjaa into 4 columns
  mutate(vj1 = str_split_i(bulk_chain, ";", 1),
         vj2 = str_split_i(bulk_chain, ";", 2),
         aa1 = str_split_i(bulk_chain, ";", 3),
         aa2 = str_split_i(bulk_chain, ";", 4),
         vjaa1 = paste0(vj1, ";", aa1),
         vjaa2 = paste0(vj2, ";", aa2)) %>%
  select(Patient, tumor_vjaa_counts, vjaa1, vjaa2, vjaa) %>%
  pivot_longer(cols = c("vjaa1", "vjaa2"), names_to = "chain", values_to = "bulk_chain") %>%
  select(-chain)

sctils_clone_chains <- rbind(sctils_chain_one, sctils_chain_multi)

# Check that all clones are unique across patients
sctils_clone_chains %>% distinct(vjaa, Patient, .keep_all = TRUE) %>% filter(duplicated(vjaa))
```

Skin
```{r}
# Splitting TRA and TRB from one clone into separate rows
scskin_clone_chains <- sc_skin@meta.data %>%
  group_by(Patient) %>%
  dplyr::count(vjaa, Patient) %>%
  arrange(desc(n)) %>%
  dplyr::rename("skin_vjaa_counts" = "n") %>%
  mutate(bulk_chain = strsplit(as.character(vjaa), "_")) %>%
  unnest(bulk_chain) %>%
  filter(bulk_chain != "NA;NA")

# Get clones with just 1 TRA or TRB chain
scskin_chain_one <- scskin_clone_chains %>%
  filter(str_count(bulk_chain, ";") == 1)

# Unnest clones with more than 1 TRA or TRB chain
scskin_chain_multi <- scskin_clone_chains %>%
  filter(str_count(bulk_chain, ";") > 1) %>%
  # split vjaa into 4 columns
  mutate(vj1 = str_split_i(bulk_chain, ";", 1),
         vj2 = str_split_i(bulk_chain, ";", 2),
         aa1 = str_split_i(bulk_chain, ";", 3),
         aa2 = str_split_i(bulk_chain, ";", 4),
         vjaa1 = paste0(vj1, ";", aa1),
         vjaa2 = paste0(vj2, ";", aa2)) %>%
  select(Patient, skin_vjaa_counts, vjaa1, vjaa2, vjaa) %>%
  pivot_longer(cols = c("vjaa1", "vjaa2"), names_to = "chain", values_to = "bulk_chain") %>%
  select(-chain)

scskin_clone_chains <- rbind(scskin_chain_one, scskin_chain_multi)

# All clones are unique across patients
scskin_clone_chains %>% distinct(vjaa, Patient, .keep_all = TRUE) %>% filter(duplicated(vjaa))
```

## Merge clones into one dataframe and count clone frequency over both tumor and skin
```{r}
sc_clone_chains <- sctils_clone_chains %>%
  full_join(scskin_clone_chains, by = c("Patient", "vjaa", "bulk_chain")) %>%
  replace(is.na(.), 0) %>%
  mutate(total_vjaa_counts = tumor_vjaa_counts + skin_vjaa_counts) %>%
  arrange(desc(total_vjaa_counts))
```

## Create clone IDs based on decreasing clone frequency in both skin and tumor
```{r}
sc_clone_ids <- sc_clone_chains %>%
  group_by(Patient) %>%
  distinct(Patient, vjaa) %>%
  mutate(obs = 1:n(),
         sc_clone_id = paste0(Patient, "_", obs)) %>%
  select(-obs) 
```

## Add clone IDs to full dataframe
```{r}
sc_clone_chains <- left_join(sc_clone_chains, sc_clone_ids, by = c("Patient", "vjaa")) %>%
  select(Patient, sc_clone_id, vjaa, bulk_chain, tumor_vjaa_counts, skin_vjaa_counts, total_vjaa_counts)
```

## Reformat chains to match bulk format
```{r}
sc_clone_chains <- sc_clone_chains %>% 
  # Reformat certain V genes to match bulk (bulk can't differentiate btwn certain V genes)
  mutate(bulk_chain = str_replace_all(bulk_chain, "TRBV12-3|TRBV12-4", "TRBV12-3/4"),
         bulk_chain = str_replace_all(bulk_chain, "TRBV6-2|TRBV6-3|TRBV6-5|TRBV6-6", "TRBV6-2/3/5/6"),
         bulk_chain = str_replace_all(bulk_chain, "TRAV29/DV5", "TRAV29DV5"),
         bulk_chain = str_replace_all(bulk_chain, "TRAV14/DV4", "TRAV14DV4"),
         bulk_chain = str_replace_all(bulk_chain, "TRAV23/DV6", "TRAV23DV6"),
         bulk_chain = str_replace_all(bulk_chain, "TRAV36/DV7", "TRAV36DV7"),
         bulk_chain = str_replace_all(bulk_chain, "TRAV38-2/DV8", "TRAV38-2DV8")) %>%
  # Reformat the . separator to ; as per bulk
  mutate(bulk_chain = str_replace_all(bulk_chain, fixed("."), ";"),)

```

## Add scclone ID and tumor, skin and total counts to scRNA objects
```{r}
# Make sure there's barcode names
sc_skin$barcodes <- rownames(sc_skin@meta.data)
sc_tils$barcodes <- rownames(sc_tils@meta.data)

sc_clones_unique <- sc_clone_chains %>%
  select(Patient, vjaa, skin_vjaa_counts, tumor_vjaa_counts, sc_clone_id) %>%
  distinct(.keep_all = TRUE)

sc_skin@meta.data <- sc_skin@meta.data %>%
  left_join(sc_clones_unique, by = c("Patient", "vjaa"))
rownames(sc_skin@meta.data) <- sc_skin$barcodes

sc_tils@meta.data <- sc_tils@meta.data %>%
  left_join(sc_clones_unique, by = c("Patient", "vjaa"))
rownames(sc_tils@meta.data) <- sc_tils$barcodes
```

## Save scRNA objects
```{r}
saveRDS(sc_tils, "scRNA_TCR_Tumor_Part4.Rds")

saveRDS(sc_skin, "scRNA_TCR_Skin_Part4.Rds")
```

## Save clones
```{r}
write.csv(sc_clone_chains, "scRNA_TCR_bulkTCR_chain_mapping_Part4.csv", row.names = FALSE)
```

## Get session info
```{r}
sessionInfo()
```

