# scTCRseq: Plot significantly expanded (PBMC-derived) and scTIL beta chain overlap {sec-sctcr-significantly-expanded-venn-diagram}

## Set up workspace
```{r}
# Libraries
library(tidyverse)
library(Seurat)
library(dplyr)
library(scRepertoire)
library(Vennerable)
library(grid)
```

## Open mapping of single cell clonotypes to separated alpha and beta chains from scTCR data (regardless of scRNA)
```{r}
sc_bulk_mapping <- read.csv("scTCR_bulkTCR_chain_mapping_Part4.csv")
```

## Load categories
```{r}
p101_til_categories <- read.csv("sctcr_scRep_p101_til_freq_Part2.csv") %>%
  mutate(Patient = "P101")
p103_til_categories <- read.csv("sctcr_scRep_p103_til_freq_Part2.csv") %>%
  mutate(Patient = "P103")

til_categories <- rbind(p101_til_categories[,c("Patient", "vjaa", "sctcr_category")], p103_til_categories[,c("Patient", "vjaa", "sctcr_category")])
```

## Join categories
```{r}
sc_bulk_mapping <- sc_bulk_mapping %>%
  left_join(til_categories) %>%
  filter(tumor_vjaa_counts > 0)

sc_bulk_mapping %>%
  group_by(Patient, sctcr_category) %>%
  dplyr::count()
```

## Load typed clones 
```{r}
p101_betas_typed <- read.csv("p101_betas_merged_typed_Part3.csv") %>%
  mutate(Beta_clonotype_Pt = paste0(Patient, "_", Beta_clonotype)) %>%
  dplyr::rename("bulk_Type" = "Type")
p103_betas_typed <- read.csv("p103_betas_merged_typed_Part3.csv") %>%
  mutate(Beta_clonotype_Pt = paste0(Patient, "_", Beta_clonotype)) %>%
  dplyr::rename("bulk_Type" = "Type")
```

# Load clones that significantly expanded after nivolumab and vaccination 
```{r}
p101_se_nivo <- read.csv("p101_pval_prevax_vs_pretreatment_Part4.csv") %>%
  filter(sig == "Sig") %>%
  # Select important columns
    select(Beta_clonotype, sig, Patient) %>%
    # Rename sig
    dplyr::rename("SE_nivo" = "sig")
p101_se_vax <- read.csv("p101_pval_postvax_vs_prevax_Part4.csv") %>%
  filter(sig == "Sig") %>%
  mutate(Patient = "P101") %>%
    select(Beta_clonotype, sig, Patient) %>%
    dplyr::rename("SE_vax" = "sig")

p103_se_nivo <- read.csv("p103_pval_prevax_vs_pretreatment_Part4.csv") %>%
  filter(sig == "Sig") %>%
    select(Beta_clonotype, sig, Patient) %>%
    dplyr::rename("SE_nivo" = "sig")
p103_se_vax <- read.csv("p103_pval_postvax_vs_prevax_Part4.csv") %>%
  filter(sig == "Sig") %>%
  mutate(Patient = "P103") %>%
    select(Beta_clonotype, sig, Patient) %>%
    dplyr::rename("SE_vax" = "sig")
```

## Set up data
```{r}
# P101
p101_w00_sctils_betas <- sc_bulk_mapping %>%
  filter(str_detect(bulk_chain, "TRB"),
         Patient %in% c("P101"),
         sctcr_category == "Existing")
p101_w12_sctils_betas <- sc_bulk_mapping %>%
  filter(str_detect(bulk_chain, "TRB"),
         Patient %in% c("P101"),
         sctcr_category == "Post-Nivolumab")
p101_w20_sctils_betas <- sc_bulk_mapping %>%
  filter(str_detect(bulk_chain, "TRB"),
         Patient %in% c("P101"),
         sctcr_category == "Post-Vaccine")

# P103
p103_w00_sctils_betas <- sc_bulk_mapping %>%
  filter(str_detect(bulk_chain, "TRB"),
         Patient %in% c("P103"),
         sctcr_category == "Existing")
p103_w12_sctils_betas <- sc_bulk_mapping %>%
  filter(str_detect(bulk_chain, "TRB"),
         Patient %in% c("P103"),
         sctcr_category == "Post-Nivolumab")
p103_w20_sctils_betas <- sc_bulk_mapping %>%
  filter(str_detect(bulk_chain, "TRB"),
         Patient %in% c("P103"),
         sctcr_category == "Post-Vaccine")
```

## Plot the overlap of significantly expanding clones after nivo, after vaccine, and the scTIL beta chains at each timepoint for Fig S7G
### P101 W00
```{r}
p101_w00_dat <- Venn(list(`Pre-treatment scTILs` = p101_w00_sctils_betas$bulk_chain, 
                           `SE Vax` = p101_se_vax$Beta_clonotype,
                           `SE Nivo` = p101_se_nivo$Beta_clonotype))

# Compute venn object
p101_w00_venn <- compute.Venn(p101_w00_dat, doWeights = TRUE)

# Create graphical params
p101_gp <- VennThemes(p101_w00_venn)

# Adjust graphical params
# Adjust the color of the circles borders
p101_gp[["Set"]][["Set1"]]$col <- "#9f2dea"
p101_gp[["Set"]][["Set2"]]$col <- "#0072B2"
p101_gp[["Set"]][["Set3"]]$col <- "darkgoldenrod3"
# p101_gp[["Set"]][["Set1"]]$col <- "black" 
# p101_gp[["Set"]][["Set2"]]$col <- "black" 
# p101_gp[["Set"]][["Set3"]]$col <- "black" 
# Adjust the color of the circles themselves
p101_gp[["Face"]][["110"]]$fill <- "#7161e499"
p101_gp[["Face"]][["101"]]$fill <- "#eb6f6f99"
p101_gp[["Face"]][["100"]]$fill <- "#9f2dea90"
p101_gp[["Face"]][["010"]]$fill <- "#1659b790"
p101_gp[["Face"]][["001-1"]]$fill <- "#e5a64d90"
p101_gp[["Face"]][["000"]]$fill <- "white"
# Adjust the color of the labels
p101_gp[["SetText"]][["Set1"]]$col <- "#9f2dea"
p101_gp[["SetText"]][["Set2"]]$col <- "#0072B2"
p101_gp[["SetText"]][["Set3"]]$col <- "darkgoldenrod3"
# Adjust the size of the numbers
p101_gp[["FaceText"]][["000"]]$fontsize <- 0

grid.newpage()
plot(p101_w00_venn, gp = p101_gp)
```

## P101 W12
```{r}
p101_w12_dat <- Venn(list(`Post-Nivolumab scTILs` = p101_w12_sctils_betas$bulk_chain, 
                           `SE Vax` = p101_se_vax$Beta_clonotype,
                           `SE Nivo` = p101_se_nivo$Beta_clonotype))

# Compute venn object
p101_w12_venn <- compute.Venn(p101_w12_dat, doWeights = TRUE)

grid.newpage()
plot(p101_w12_venn, gp = p101_gp)
```

## P101 W20
```{r}
p101_w20_dat <- Venn(list(`Post-vaccine scTILs` = p101_w20_sctils_betas$bulk_chain, 
                           `SE Vax` = p101_se_vax$Beta_clonotype,
                           `SE Nivo` = p101_se_nivo$Beta_clonotype))

# Compute venn object
p101_w20_venn <- compute.Venn(p101_w20_dat, doWeights = TRUE)

grid.newpage()
plot(p101_w20_venn, gp = p101_gp)
```

## P103 W00
```{r}
p103_w00_dat <- Venn(list(`Pre-treatment scTILs` = p103_w00_sctils_betas$bulk_chain, 
                           `SE Vax` = p103_se_vax$Beta_clonotype,
                           `SE Nivo` = p103_se_nivo$Beta_clonotype))

# Compute venn object
p103_w00_venn <- compute.Venn(p103_w00_dat, doWeights = FALSE)

# Create graphical params for P103
p103_gp_00 <- VennThemes(p103_w00_venn)

# Adjust graphical params
# Adjust the color of the circles borders
p103_gp_00[["Set"]][["Set1"]]$col <- "#9f2dea"
p103_gp_00[["Set"]][["Set2"]]$col <- "#0072B2"
p103_gp_00[["Set"]][["Set3"]]$col <- "darkgoldenrod3"
# p103_gp[["Set"]][["Set1"]]$col <- "black" 
# p103_gp[["Set"]][["Set2"]]$col <- "black" 
# p103_gp[["Set"]][["Set3"]]$col <- "black" 
# Adjust the color of the circles themselves
p103_gp_00[["Face"]][["110"]]$fill <- "#7161e499"
p103_gp_00[["Face"]][["101"]]$fill <- "#eb6f6f99"
p103_gp_00[["Face"]][["100"]]$fill <- "#9f2dea90"
p103_gp_00[["Face"]][["010"]]$fill <- "#1659b790"
p103_gp_00[["Face"]][["001-1"]]$fill <- "#e5a64d90"
p103_gp_00[["Face"]][["011"]]$fill <- "#87782190"
p103_gp_00[["Face"]][["001"]]$fill <- "#ce550090"
p103_gp_00[["Face"]][["000"]]$fill <- "#6f673999"
# Adjust the color of the labels
p103_gp_00[["SetText"]][["Set1"]]$col <- "#9f2dea"
p103_gp_00[["SetText"]][["Set2"]]$col <- "#0072B2"
p103_gp_00[["SetText"]][["Set3"]]$col <- "darkgoldenrod3"
# Adjust the size of the numbers
p103_gp_00[["FaceText"]][["000"]]$fontsize <- 0

grid.newpage()
plot(p103_w00_venn, gp = p103_gp_00)
```

## P101 W12
```{r}
p103_w12_dat <- Venn(list(`Post-Nivolumab scTILs` = p103_w12_sctils_betas$bulk_chain, 
                           `SE Vax` = p103_se_vax$Beta_clonotype,
                           `SE Nivo` = p103_se_nivo$Beta_clonotype))

# Compute venn object
p103_w12_venn <- compute.Venn(p103_w12_dat, doWeights = FALSE)


# Create graphical params for P103
p103_gp <- VennThemes(p103_w12_venn)

# Adjust graphical params
# Adjust the color of the circles borders
p103_gp[["Set"]][["Set1"]]$col <- "#9f2dea"
p103_gp[["Set"]][["Set2"]]$col <- "#0072B2"
p103_gp[["Set"]][["Set3"]]$col <- "darkgoldenrod3"
# p103_gp[["Set"]][["Set1"]]$col <- "black" 
# p103_gp[["Set"]][["Set2"]]$col <- "black" 
# p103_gp[["Set"]][["Set3"]]$col <- "black" 
# Adjust the color of the circles themselves
p103_gp[["Face"]][["110"]]$fill <- "#7161e499"
p103_gp[["Face"]][["101"]]$fill <- "#eb6f6f99"
p103_gp[["Face"]][["100"]]$fill <- "#9f2dea90"
p103_gp[["Face"]][["010"]]$fill <- "#1659b790"
p103_gp[["Face"]][["011"]]$fill <- "#ce550090"
p103_gp[["Face"]][["001"]]$fill <- "#e5a64d90"
p103_gp[["Face"]][["000"]]$fill <- "#6f673999"
# Adjust the color of the labels
p103_gp[["SetText"]][["Set1"]]$col <- "#9f2dea"
p103_gp[["SetText"]][["Set2"]]$col <- "#0072B2"
p103_gp[["SetText"]][["Set3"]]$col <- "darkgoldenrod3"

grid.newpage()
plot(p103_w12_venn, gp = p103_gp)
```

## P103 W20
```{r}
p103_w20_dat <- Venn(list(`Post-vaccine scTILs` = p103_w20_sctils_betas$bulk_chain, 
                           `SE Vax` = p103_se_vax$Beta_clonotype,
                           `SE Nivo` = p103_se_nivo$Beta_clonotype))

# Compute venn object
p103_w20_venn <- compute.Venn(p103_w20_dat, doWeights = FALSE)

grid.newpage()
plot(p103_w20_venn, gp = p103_gp)
```

## Get session info
```{r}
sessionInfo()
```
