# scTCRseq: Compile and format data {sec-sctcr-compile-and-format-data}

Some scRNA and scTCR tumor timepoints were relabelled after this analysis was conducted. These are the corrected timepoints:
101: W20 -> W24
103: W20 -> W25

## Set up workspace 
```{r}
# Load libraries
library(tidyverse)
library(Seurat)
library(dplyr)
library(scRepertoire)
```

Functions to create the custom definition of clones
```{r}
# Turn all entries of c_gene into ""
to_empty <- function(df){
  df$c_gene <- ""
  return(df)
}

# Count number of TRA and TRB per barcode
count_TRA_TRB <- function(df){
  df <- df %>%
    mutate(TCR1_count = str_count(TCR1, fixed(";")) + 1,
           TCR2_count = str_count(TCR2, fixed(";")) + 1,
           TCR1_count = replace_na(TCR1_count, 0),
           TCR2_count = replace_na(TCR2_count, 0))
}

# Remove the barcode if it has 2A+2B or 3A/XB or XA/3B
remove_multi_chains <- function(df){
  df <- df %>%
    # 2A+2B
    filter(!(TCR1_count == 2 & TCR2_count == 2),
           !(TCR1_count >= 3),
           !(TCR2_count >= 3))
}

# Remove "NA" if it was supposed to be the "C" gene in TRA
rm_na_TCR1 <- function(df){
  df <- df %>%
    mutate(TCR1 = str_replace_all(TCR1, "(\\.NA(?=;))|(\\.NA$)", ""))
  return(df)
}

# Remove "NA" if it was supposed to be the "C" or "D" gene in TRB (D gene is not seen in bulkTCR data anyway)
rm_na_TCR2 <- function(df){
  df <- df %>%
    mutate(TCR2 = str_replace_all(TCR2, "(\\.NA(?=;))|(\\.NA(?=.))|(\\.NA$)", ""))
  return(df)
}


# Create new columns: CTVJaa, CTbeta, CTalpha
create_columns <- function(df){
  df <- df %>%
    mutate(vjaa = paste0(TCR1, ";", cdr3_aa1, "_", TCR2, ";", cdr3_aa2),
           alpha = paste0(TCR1, ";", cdr3_aa1),
           beta = paste0(TCR2, ";", cdr3_aa2))
}

```


## Load scTCR clones
```{r}
P101_Skin_Post3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P101_Skin_Post3rd_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P101_Skin_Pre3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P101_Skin_Pre3rd_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P101_Tumor_W00 <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P101_Tumor_W00_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P101_Tumor_W12 <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P101_Tumor_W12_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P101_Tumor_W20 <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P101_Tumor_W20_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P103_Skin_Post3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P103_Skin_Post3rd_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P103_Skin_Pre3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P103_Skin_Pre3rd_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P103_Tumor_W00 <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P103_Tumor_W00_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P103_Tumor_W12 <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P103_Tumor_W12_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P103_Tumor_W20 <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P103_Tumor_W20_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P104_Skin_Post3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P104_Skin_Post3rd_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P104_Skin_Pre3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P104_Skin_Pre3rd_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P104_Tumor_PD <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P104_Tumor_PD_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P105_Skin_Post3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P105_Skin_Post3rd_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P105_Skin_Pre3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P105_Skin_Pre3rd_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P106_Skin_Post3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P106_Skin_Post3rd_2-5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P106_Skin_Pre3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P106_Skin_Pre3rd_2-5mgIpi_TCR_fixed021324_cellranger/outs/filtered_contig_annotations.csv")
P108_Skin_Post3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P108_Skin_Post3rd_5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P108_Skin_Pre3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P108_Skin_Pre3rd_5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P108_Tumor_PD <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P108_Tumor_PD_5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P109_Skin_Pre3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P109_Skin_Pre3rd_5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P110_Skin_Post3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P110_Skin_Post3rd_5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P110_Skin_Pre3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P110_Skin_Pre3rd_5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P111_Skin_Post3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P111_Skin_Post3rd_5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
P111_Skin_Pre3rd <- read.csv("/jsimonlab/projects/Wu/Melanoma_scTCR_Eryn/melanoma_sctcr_reprocessed_040224/P111_Skin_Pre3rd_5mgIpi_TCR_cellranger/outs/filtered_contig_annotations.csv")
```


## Combine TCRs and define clones using the V + J gene + cdr3AA
```{r}
contig_list <- list(P101_Skin_Post3rd, P101_Skin_Pre3rd, P101_Tumor_W00, P101_Tumor_W12, P101_Tumor_W20, P103_Skin_Post3rd, P103_Skin_Pre3rd, P103_Tumor_W00, P103_Tumor_W12, P103_Tumor_W20, P104_Skin_Post3rd, P104_Skin_Pre3rd, P104_Tumor_PD, P105_Skin_Post3rd, P105_Skin_Pre3rd, P106_Skin_Post3rd, P106_Skin_Pre3rd, P108_Skin_Post3rd, P108_Skin_Pre3rd, P108_Tumor_PD, P109_Skin_Pre3rd, P110_Skin_Post3rd, P110_Skin_Pre3rd, P111_Skin_Post3rd, P111_Skin_Pre3rd)
sample_list <- c("P101_Skin_Post3rd", "P101_Skin_Pre3rd", "P101_Tumor_W00", "P101_Tumor_W12", "P101_Tumor_W20", "P103_Skin_Post3rd", "P103_Skin_Pre3rd", "P103_Tumor_W00", "P103_Tumor_W12", "P103_Tumor_W20", "P104_Skin_Post3rd", "P104_Skin_Pre3rd", "P104_Tumor_PD", "P105_Skin_Post3rd", "P105_Skin_Pre3rd", "P106_Skin_Post3rd", "P106_Skin_Pre3rd", "P108_Skin_Post3rd", "P108_Skin_Pre3rd", "P108_Tumor_PD", "P109_Skin_Pre3rd", "P110_Skin_Post3rd", "P110_Skin_Pre3rd", "P111_Skin_Post3rd", "P111_Skin_Pre3rd")

# Create a new column with the combination of the amino acid sequence and gene sequence, without the C gene. ie: VJcdr3AA

contig_list <- lapply(contig_list, to_empty)

combined_TCR <- combineTCR(contig_list, 
                           samples = sample_list,
                           removeNA = FALSE, 
                           removeMulti = FALSE, 
                           filterMulti = FALSE)

# Apply counts and filter
combined_TCR <- lapply(combined_TCR, count_TRA_TRB)

# Supplementary table creation
# Count the number of clones that have less than 2A/2B or 3A/B
suppTable <- do.call(rbind, combined_TCR)

suppTable %>%
  group_by(sample) %>%
    mutate(to_remove = case_when((TCR1_count == 2 & TCR2_count == 2) ~ "MultiChain",
                                 (TCR1_count >= 3) ~ "MultiChain",
                                 (TCR2_count >= 3) ~ "MultiChain",
                                 T ~ "SingleChain")) %>%
  dplyr::count(to_remove) %>%
  pivot_wider(names_from = "to_remove", values_from = "n")

combined_TCR <- lapply(combined_TCR, remove_multi_chains)

# Apply removals
combined_TCR <- lapply(combined_TCR, rm_na_TCR1)
combined_TCR <- lapply(combined_TCR, rm_na_TCR2)

# Apply new columns
combined_TCR <- lapply(combined_TCR, create_columns)

```

## Add metadata
```{r}
combined_TCR <- addVariable(combined_TCR, 
                            variable.name = "Patient", 
                            variables = str_replace_all(sample_list,"_.+",""))
combined_TCR <- addVariable(combined_TCR, 
                            variable.name = "Site", 
                            variables = str_replace_all(sample_list,".+_(.+)_.+","\\1"))
combined_TCR <- addVariable(combined_TCR, 
                            variable.name = "Timepoint", 
                            variables = str_replace_all(sample_list,".+_",""))
```

## Save data with my custom definition of clones in the "vjaa" column
```{r}
saveRDS(combined_TCR, "sctcr_scRep_combined_TCR_skin_tumor_Part1.Rds")
```

## Get session info
```{r}
sessionInfo()
```

