# scTCRseq Skin: Attach scTCR and categorizations to scRNA object {sec-skin-attach-sctcr-scrna}

## Set up workspace
```{r}
# Libraries
library(tidyverse)
library(Seurat)
library(dplyr)
library(scRepertoire)
library(data.table)
```

## Load scRepertoire skin and tumor data with my custom vjaa column
```{r}
combined_TCR <- readRDS("sctcr_scRep_combined_TCR_skin_tumor_Part1.Rds")
```

## Load categorizations of scTCR clones
```{r}
p101_skin <- read.csv("sctcr_scRep_p101_skin_freq_Part2.csv")
p103_skin <- read.csv("sctcr_scRep_p103_skin_freq_Part2.csv")
p104_skin <- read.csv("sctcr_scRep_p104_skin_freq_Part2.csv")
p105_skin <- read.csv("sctcr_scRep_p105_skin_freq_Part2.csv")
p106_skin <- read.csv("sctcr_scRep_p106_skin_freq_Part2.csv")
p108_skin <- read.csv("sctcr_scRep_p108_skin_freq_Part2.csv")
p109_skin <- read.csv("sctcr_scRep_p109_skin_freq_Part2.csv")
p110_skin <- read.csv("sctcr_scRep_p110_skin_freq_Part2.csv")
p111_skin <- read.csv("sctcr_scRep_p111_skin_freq_Part2.csv")
```

## Reformat dataframes so that each scTCR barcode has its own row
```{r}
p101_skin_reformatted <- data.frame()
for(i in c(1:nrow(p101_skin))){
  tcr_categories <- p101_skin[i, c("sctcr_category", "sctcr_category2", "sctcr_category3", "reactive")]
  barcodes <- p101_skin[i,"barcodes"]
  barcodes <- str_split_1(barcodes, ",")
  reformatted_df <- as.data.frame(barcodes)
  reformatted_df <- cbind(reformatted_df, tcr_categories)
  p101_skin_reformatted <- rbind(p101_skin_reformatted, reformatted_df)
}
rownames(p101_skin_reformatted) <- p101_skin_reformatted$barcodes
  
p103_skin_reformatted <- data.frame()
for(i in c(1:nrow(p103_skin))){
  tcr_categories <- p103_skin[i, c("sctcr_category", "sctcr_category2", "sctcr_category3", "reactive")]
  barcodes <- p103_skin[i,"barcodes"]
  barcodes <- str_split_1(barcodes, ",")
  reformatted_df <- as.data.frame(barcodes)
  reformatted_df <- cbind(reformatted_df, tcr_categories)
  p103_skin_reformatted <- rbind(p103_skin_reformatted, reformatted_df)
}
rownames(p103_skin_reformatted) <- p103_skin_reformatted$barcodes

p104_skin_reformatted <- data.frame()
for(i in c(1:nrow(p104_skin))){
  tcr_categories <- p104_skin[i, c("sctcr_category", "sctcr_category2", "sctcr_category3", "reactive")]
  barcodes <- p104_skin[i,"barcodes"]
  barcodes <- str_split_1(barcodes, ",")
  reformatted_df <- as.data.frame(barcodes)
  reformatted_df <- cbind(reformatted_df, tcr_categories)
  p104_skin_reformatted <- rbind(p104_skin_reformatted, reformatted_df)
}
rownames(p104_skin_reformatted) <- p104_skin_reformatted$barcodes

p105_skin_reformatted <- data.frame()
for(i in c(1:nrow(p105_skin))){
  tcr_categories <- p105_skin[i, c("sctcr_category", "sctcr_category2", "sctcr_category3", "reactive")]
  barcodes <- p105_skin[i,"barcodes"]
  barcodes <- str_split_1(barcodes, ",")
  reformatted_df <- as.data.frame(barcodes)
  reformatted_df <- cbind(reformatted_df, tcr_categories)
  p105_skin_reformatted <- rbind(p105_skin_reformatted, reformatted_df)
}
rownames(p105_skin_reformatted) <- p105_skin_reformatted$barcodes

p106_skin_reformatted <- data.frame()
for(i in c(1:nrow(p106_skin))){
  tcr_categories <- p106_skin[i, c("sctcr_category", "sctcr_category2", "sctcr_category3", "reactive")]
  barcodes <- p106_skin[i,"barcodes"]
  barcodes <- str_split_1(barcodes, ",")
  reformatted_df <- as.data.frame(barcodes)
  reformatted_df <- cbind(reformatted_df, tcr_categories)
  p106_skin_reformatted <- rbind(p106_skin_reformatted, reformatted_df)
}
rownames(p106_skin_reformatted) <- p106_skin_reformatted$barcodes

p108_skin_reformatted <- data.frame()
for(i in c(1:nrow(p108_skin))){
  tcr_categories <- p108_skin[i, c("sctcr_category", "sctcr_category2", "sctcr_category3", "reactive")]
  barcodes <- p108_skin[i,"barcodes"]
  barcodes <- str_split_1(barcodes, ",")
  reformatted_df <- as.data.frame(barcodes)
  reformatted_df <- cbind(reformatted_df, tcr_categories)
  p108_skin_reformatted <- rbind(p108_skin_reformatted, reformatted_df)
}
rownames(p108_skin_reformatted) <- p108_skin_reformatted$barcodes

p109_skin_reformatted <- data.frame()
for(i in c(1:nrow(p109_skin))){
  tcr_categories <- p109_skin[i, c("sctcr_category", "sctcr_category2", "sctcr_category3", "reactive")]
  barcodes <- p109_skin[i,"barcodes"]
  barcodes <- str_split_1(barcodes, ",")
  reformatted_df <- as.data.frame(barcodes)
  reformatted_df <- cbind(reformatted_df, tcr_categories)
  p109_skin_reformatted <- rbind(p109_skin_reformatted, reformatted_df)
}
rownames(p109_skin_reformatted) <- p109_skin_reformatted$barcodes

p110_skin_reformatted <- data.frame()
for(i in c(1:nrow(p110_skin))){
  tcr_categories <- p110_skin[i, c("sctcr_category", "sctcr_category2", "sctcr_category3", "reactive")]
  barcodes <- p110_skin[i,"barcodes"]
  barcodes <- str_split_1(barcodes, ",")
  reformatted_df <- as.data.frame(barcodes)
  reformatted_df <- cbind(reformatted_df, tcr_categories)
  p110_skin_reformatted <- rbind(p110_skin_reformatted, reformatted_df)
}
rownames(p110_skin_reformatted) <- p110_skin_reformatted$barcodes

p111_skin_reformatted <- data.frame()
for(i in c(1:nrow(p111_skin))){
  tcr_categories <- p111_skin[i, c("sctcr_category", "sctcr_category2", "sctcr_category3", "reactive")]
  barcodes <- p111_skin[i,"barcodes"]
  barcodes <- str_split_1(barcodes, ",")
  reformatted_df <- as.data.frame(barcodes)
  reformatted_df <- cbind(reformatted_df, tcr_categories)
  p111_skin_reformatted <- rbind(p111_skin_reformatted, reformatted_df)
}
rownames(p111_skin_reformatted) <- p111_skin_reformatted$barcodes
```

## Merge categories and reactivity per barcode into one dataframe
```{r}
skin_category_reactivity <- rbindlist(list(p101_skin_reformatted, p103_skin_reformatted, p104_skin_reformatted, p105_skin_reformatted, p106_skin_reformatted, p108_skin_reformatted, p109_skin_reformatted, p110_skin_reformatted, p111_skin_reformatted), fill = TRUE)

# Add barcode
rownames(skin_category_reactivity) <- skin_category_reactivity$barcode
```

## Open latest scRNA object
```{r}
skin.18279 <- readRDS("/jsimonlab/projects/Wu/Melanoma_scRNA_Eryn/Analysis_072324_Skin/Melanoma_18-279_SKIN_scRNA_USA_alevin-fry_seurat_integrated_harmony_clustered_withSexes_TNKsubClustered_withProjecTILs_CD4_CD8.rds")
```

## Link scTCR data to scRNA data using custom column
```{r}
combined_TCR$P101_Skin_Post3rd$barcode[1:10]
colnames(skin.18279)[1:10]

colnames(skin.18279) <- str_replace_all(str_replace_all(colnames(skin.18279),"_.{0,2}5mgIpi_RNA",""),"$","-1")
colnames(skin.18279)[1:10]

# Bugs out when you use capital letters in your custom call
skin.18279 <- combineExpression(combined_TCR, 
                         skin.18279, 
                         cloneCall="vjaa", 
                         group.by = "sample", 
                         proportion = TRUE)
```

## Attach categories into metadata
```{r}
skin.18279 <- AddMetaData(skin.18279, skin_category_reactivity)
```

## Save scRNA object with all TCRs formatted via scRep and TCR reactivities
```{r}
saveRDS(skin.18279, "scRNA_TCR_Skin_Part3.Rds")
```

## Get session info
```{r}
sessionInfo()
```

