# IFNg: Classification of CD4 and CD8 T cell states {#sec-ifng-projectils}

## Set up Seurat workspace

```{r }
#| warning: false
#| error: false

# Load libraries
library(data.table)
library(devtools)
library(presto)
library(glmGamPoi)
library(sctransform)
library(Seurat)
library(tidyverse)
library(miQC)  
library(SeuratWrappers)
library(flexmix)
library(SingleCellExperiment)
library(SummarizedExperiment)
library(readxl)
library(fishpond)
library(Matrix)
library(speckle)
library(scater)
library(patchwork)
library(vctrs)
library(alevinQC)
library(harmony)
library(scDblFinder)
library(cellXY)
library(STACAS)
library(ProjecTILs)

# Set global options for Seurat v5 objects
options(Seurat.object.assay.version = 'v5')
```

## Load previously saved clustered object

```{r}
merged.18279.singlets <- readRDS("IFNg_scRNA_Part2.rds")
```

## Run CD8 and CD4 projection with `projecTILs` as in [this vignette](https://carmonalab.github.io/ProjecTILs_CaseStudies/Bassez_BC.html)

Note these classifications are limited only to CD4/CD8 and do not include NK cells

### Retrieve human CD8 and CD4 reference maps

```{r}
#| cache: true
#| cache-lazy: false
options(timeout = max(900, getOption("timeout")))
download.file("https://figshare.com/ndownloader/files/41414556", destfile = "CD8T_human_ref_v1.rds")
ref.cd8 <- load.reference.map("CD8T_human_ref_v1.rds")

download.file("https://figshare.com/ndownloader/files/39012395", destfile = "CD4T_human_ref_v1.rds")
ref.cd4 <- load.reference.map("CD4T_human_ref_v1.rds")
```

### Plot reference atlases

```{r}
#| fig-width: 10
a <- DimPlot(ref.cd8, cols = ref.cd8@misc$atlas.palette, label = T) + theme(aspect.ratio = 1) +
    ggtitle("CD8 T reference") + NoLegend()

b <- DimPlot(ref.cd4, cols = ref.cd4@misc$atlas.palette, label = T) + theme(aspect.ratio = 1) +
    ggtitle("CD4 T reference") + NoLegend()

a | b
```

### Classify CD8 T subtypes

```{r}
merged.18279.singlets <- ProjecTILs.classifier(merged.18279.singlets, ref.cd8, ncores = 8, split.by = "Sample")
```

### Classify CD4 T subtypes

```{r}
merged.18279.singlets <- ProjecTILs.classifier(merged.18279.singlets, ref.cd4, ncores = 8, split.by = "Sample", overwrite = F)
```

## Tabulate and visualize projections

In addition to guiding annotation, these projected cell states can be quantified downstream comparing pre vs post-vaccine, tumor samples over time, or to characterize the phenotypes of T-cells that have expanded with vaccination

### Count total cells by T cell state

```{r}
table(merged.18279.singlets$functional.cluster, useNA = "ifany")
```

### Tabulate which T cell state was most abundant in each cluster at 0.7 res

```{r}
as.data.frame(table(merged.18279.singlets$RNA_snn_res.0.7, merged.18279.singlets$functional.cluster, useNA = "ifany")) %>%
    as_tibble() %>%
    dplyr::rename("Cluster" = Var1, "CellState" = Var2) %>%
    group_by(Cluster) %>%
    slice_max(Freq,n=1) %>%
    inner_join(enframe(table(merged.18279.singlets$RNA_snn_res.0.7),name="Cluster",value="TotalCount"),by="Cluster") %>%
    as.data.frame()
```

### Plot UMAP of all cells labeled by projected T cell state

```{r}
DimPlot(merged.18279.singlets, group.by = "functional.cluster", reduction="umap.harmony")
```

## Plot UMAP labeled by CD3, CD4, CD8 expression and ProjecTILs annotations
```{r}
#| fig-height: 10
#| fig-width: 10
FeaturePlot(merged.18279.singlets,
        features = c("CD3E", "CD3D", "CD8A", "CD4"),
        reduction = "umap.harmony", 
        order = TRUE) +
DimPlot(merged.18279.singlets,
        group.by = "functional.cluster",
        reduction = "umap.harmony") 
```

## Save modified object

```{r}
saveRDS(merged.18279.singlets, "IFNg_scRNA_Part3.rds")
```

## Get session info

```{r}
sessionInfo()
```
