# Skin: Post-clustering QC and characterization {#sec-skin-initial-cluster-qc}

## Set up Seurat workspace

```{r }
#| warning: false
#| error: false

# Load libraries
library(data.table)
library(devtools)
library(presto)	
library(glmGamPoi)
library(sctransform)
library(Seurat)	
library(tidyverse)
library(miQC)
library(SeuratWrappers)
library(flexmix)
library(SingleCellExperiment)
library(SummarizedExperiment)
library(readxl)
library(fishpond)
library(Matrix)
library(speckle)
library(scater)
library(patchwork)
library(vctrs)
library(alevinQC)
library(harmony)
library(scDblFinder)
library(cellXY)

# Set global options for Seurat v5 objects
options(Seurat.object.assay.version = 'v5')
```

## Load previously saved object

```{r}
merged.18279.skin.singlets <- readRDS("Skin_scRNA_Part3.rds")
```

## Set idents to preferred initial clustering resolution

```{r}
Idents(merged.18279.skin.singlets) <- merged.18279.skin.singlets$RNA_snn_res.0.8
merged.18279.skin.singlets$seurat_clusters <- merged.18279.skin.singlets$RNA_snn_res.0.8
DimPlot(merged.18279.skin.singlets, reduction="umap.harmony", label = TRUE)
```

## Plot QC

```{r}
#| fig-width: 15
VlnPlot(merged.18279.skin.singlets, features = "nCount_RNA", group.by="seurat_clusters",pt.size=0) + NoLegend()
VlnPlot(merged.18279.skin.singlets, features = "nFeature_RNA", group.by="seurat_clusters",pt.size=0) + NoLegend()
VlnPlot(merged.18279.skin.singlets, features = "percent.mt", group.by="seurat_clusters",pt.size=0) + NoLegend()
```

## Identify cursory marker genes of each cluster
Note layers were already joined in previous session

```{r}
DefaultAssay(merged.18279.skin.singlets) <- "RNA"

vargenes <- presto::wilcoxauc(merged.18279.skin.singlets, 'seurat_clusters', seurat_assay = 'RNA')
top_vargenes <- top_markers(vargenes, n = 100, auc_min = 0.5, pct_in_min = 50, pct_out_max = 50)
top_vargenes

```

## Plot genes from slide-tags preprint

Sourced from [this preprint](https://www.biorxiv.org/content/10.1101/2023.04.01.535228v1.full.pdf)

```{r}
goi <- c("PMEL","MLANA","CCL5","LEF1","FOXP3","PAX5","IGHG3","KCNMA1","ZNF366","CUX2","COL1A1","PLVAP")
VlnPlot(merged.18279.skin.singlets,features=goi,assay="RNA",layer="data",flip=T,sort=T,stack=T)
```

## Plot other known markers
Sourced from [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9700682/)

```{r}
#| fig-width: 20
#| fig-height: 12
tumor <- c("DCT","MLANA","MITF","PMEL","S100A1","TYR","APOC1")
endothelial <- c("PECAM1","VWF")
fibroblast <- c("COL3A1","COL1A1","COL1A2","LUM")
tcell <- c("FGFBP2","KLRD1","CD3E","CD3D","GZMB","XCL2","GZMH","CST7","GZMK","GZMA","IFNG","GNLY","CCL4","NKG7","CCL5","CD8A","CD8B","CTLA4","TNFRSF4","BATF","ITM2A")
mono <- c("LYZ","CD74","CD68")
bcell <- c("MS4A1","CD79A")

VlnPlot(merged.18279.skin.singlets,features=tumor,assay="RNA",layer="data",flip=T,sort=T,stack=T) +
  VlnPlot(merged.18279.skin.singlets,features=endothelial,assay="RNA",layer="data",flip=T,sort=T,stack=T) +
  VlnPlot(merged.18279.skin.singlets,features=fibroblast,assay="RNA",layer="data",flip=T,sort=T,stack=T) +
  VlnPlot(merged.18279.skin.singlets,features=tcell,assay="RNA",layer="data",flip=T,sort=T,stack=T) +
  VlnPlot(merged.18279.skin.singlets,features=mono,assay="RNA",layer="data",flip=T,sort=T,stack=T) +
  VlnPlot(merged.18279.skin.singlets,features=bcell,assay="RNA",layer="data",flip=T,sort=T,stack=T) +
  NoLegend()
```

## Plot top markers identified and canonical genes as a dotplot

```{r}
#| fig-width: 30
#| fig-height: 8
top_vargenes <- top_markers(vargenes, n = 5, auc_min = 0.5, pct_in_min = 50, pct_out_max = 50)
top_markers <- top_vargenes %>%
	select(-rank) %>% 
	unclass() %>% 
	stack() %>%
	pull(values) %>%
	unique() %>%
	.[!is.na(.)]

dotplotmarkers <- unique(c(top_markers,tumor,endothelial,fibroblast,tcell,mono,bcell))

# Compute aggregated expression values of these genes and cluster them to order the figure
rna <- AverageExpression(merged.18279.skin.singlets,assay="RNA",slot="data")
rna.sub <- rna$RNA[dotplotmarkers,]
cors.genes <- as.dist(1-cor(as.matrix(t(rna.sub)),method="pearson"))
hc.genes <- hclust(cors.genes)
dotplotmarkers.sorted <- rownames(rna.sub)[hc.genes$order]

# Plot
DotPlot(merged.18279.skin.singlets,features=dotplotmarkers.sorted,assay="RNA",cols=c("blue","red"),cluster.idents=T) + RotatedAxis()

```

## Plot expression of key genes in UMAP space

```{r}
#| fig-height: 16
FeaturePlot(merged.18279.skin.singlets, 
	reduction="umap.harmony", 
	features=c("CD8A","CD4","CTLA4","KLRC1","CD79A","PMEL","MLANA","LYZ","PECAM1","COL3A1","SFN","KRT19"),
	order = T,
	ncol = 2)
```

## Get session info

```{r}
sessionInfo()
```
