# Skin: Differential expression analysis, Post vs Pre 3rd vaccination {#sec-skin-differential-expression}

## Set up Seurat workspace

```{r }
#| warning: false
#| error: false

# Load libraries
library(presto)
library(Seurat)
library(tidyverse)
library(patchwork)
library(paletteer)
library(msigdbr)
library(ComplexHeatmap)
library(circlize)
library(muscat)
library(edgeR)
library(SingleCellExperiment)
library(DESeq2)
library(scuttle)
library(gprofiler2)
library(msigdbr)
```

## Load previous seurat object

```{r}
merged.18279.skin.singlets <- readRDS("Skin_scRNA_Part8.rds")
```

## Run pseudobulk differential expression test with `muscat`

### Convert Seurat object to SingleCellExperiment

```{r}
merged.18279.skin.singlets[['RNA']] <- JoinLayers(merged.18279.skin.singlets[['RNA']])
merged.sce <- as.SingleCellExperiment(merged.18279.skin.singlets, assay="RNA")
```

### Prep coldata and create pseudobulk of raw counts by sub-cluster x sample

```{r}
(mergedCondition.sce <- prepSCE(merged.sce,
	kid = "sub.cluster",
	gid = "Timepoint",
	sid = "Sample",
	drop=TRUE))

pb <- aggregateData(mergedCondition.sce,
    assay = "counts",fun="sum",
    by = c("cluster_id", "sample_id"))

```

### Construct design & contrast matrix to compare Post3rd to Pre3rd timepoints

```{r}
ei <- metadata(mergedCondition.sce)$experiment_info
mm <- model.matrix(~0 + ei$group_id)
dim_name <- levels(ei$group_id)
dimnames(mm) <- list(ei$sample_id, dim_name)
contrast <- makeContrasts("Post3rd-Pre3rd", levels = mm)
```

### Run differential expression test via DESeq2

```{r}
#| cache: true
#| cache-lazy: false
res <- pbDS(pb, 
            design = mm, 
            contrast = contrast, 
            method="DESeq2", 
            min_cells = 20,
            verbose = FALSE)

# Combine results from all clusters into one tibble
df <- bind_rows(res$table$`Post3rd-Pre3rd`, .id = "sub.cluster") %>% 
	as_tibble()

```

### Show snippet of results from one cluster: upregulated genes in cluster 7

```{r}
df %>% 
	dplyr::filter(p_adj.loc < 0.05 & sub.cluster=="7" & logFC > 0) %>% 
	dplyr::arrange(p_adj.loc) %>% 
	print(n=30)
```

### Write differential expression results to tsv

```{r}
write_tsv(df, "Skin_scRNA_PostVsPre3rd_DiffExp_results.tsv")
```

### Plot number of significant genes for all clusters

```{r}
all_subclusts <- enframe(unique(merged.18279.skin.singlets$sub.cluster),name=NULL,value="sub.cluster")
df %>%
  dplyr::filter(p_adj.loc < 0.05) %>%
  group_by(sub.cluster) %>%
  summarize(Up = count(logFC > 0),Down = (-1 * count(logFC < 0)),Total = (count(logFC > 0) + count(logFC < 0))) %>%
  left_join(all_subclusts, ., by="sub.cluster") %>%
  mutate(Up = replace_na(Up, 0), Down = replace_na(Down, 0), Total = replace_na(Total, 0)) %>%
  pivot_longer(cols = c(Up,Down), names_to="Direction", values_to="NumberDE") %>%
  ggplot(aes(x = reorder(sub.cluster,-Total), y = NumberDE, fill = Direction)) +
  geom_col() +
  theme_bw() +
  xlab("Sub-cluster") +
  ylab("Number of Differentially Expressed Genes\np_adj.loc < 0.05") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
```

## Get session info

```{r}
sessionInfo()
```
