# Skin: Check patient sexes {#sec-skin-patient-sex}

## Set up Seurat workspace

```{r }
#| warning: false
#| error: false

# Load libraries
library(data.table)
library(devtools)
library(presto)
library(glmGamPoi)
library(sctransform)
library(Seurat)	
library(tidyverse)
library(miQC)	
library(SeuratWrappers)
library(flexmix)
library(SingleCellExperiment)
library(SummarizedExperiment)
library(readxl)
library(fishpond)
library(Matrix)
library(speckle)
library(scater)
library(patchwork)
library(vctrs)
library(alevinQC)
library(harmony)
library(scDblFinder)
library(cellXY)

# Set global options for Seurat v5 objects
options(Seurat.object.assay.version = 'v5')
```

## Load previously saved clustered object

```{r}
merged.18279.skin.singlets <- readRDS("Skin_scRNA_Part2.rds")
```

## Infer sex per cell using chrX and chrY gene counts using `cellXY`

```{r}
merged.18279.skin.singlets[['RNA']] <- JoinLayers(merged.18279.skin.singlets[['RNA']])
xyPredict <- classifySex(x=merged.18279.skin.singlets@assays$RNA$counts,genome="Hs", qc=F)
merged.18279.skin.singlets <- AddMetaData(merged.18279.skin.singlets,xyPredict$prediction,col.name="CellPredictedSex")

```

## Assign known sexes to each sample

```{r}
knownSex <- as.data.frame(cbind("Sample" = unique(merged.18279.skin.singlets$Sample), "Sex" = c(rep("Male",2),rep("Male",2),rep("Female",2),rep("Male",2),rep("Female",2),rep("Male",2),rep("Female",1),rep("Male",2),rep("Male",2))))

knownSex
```

## Summarize cell-wise sex predictions per Sample and compare to known labels

If more than 80% of the individual cell sex predictions are consistent for a given Sample, we call that Sample as that sex, then match to known labels

```{r}
# Show snippet first
rownames_to_column(merged.18279.skin.singlets@meta.data,var="bc") %>%
    as_tibble() %>%
    dplyr::select(bc,CellPredictedSex,Sample) %>%
    group_by(Sample) %>%
    summarize(nMale = sum(CellPredictedSex=="Male"), 
    	nFemale = sum(CellPredictedSex=="Female"), 
    	nCells = n()
    )

# Now determine sex per sample and count matching vs non-matching labels
rownames_to_column(merged.18279.skin.singlets@meta.data,var="bc") %>%
    as_tibble() %>%
    dplyr::select(bc,CellPredictedSex,Sample) %>%
    group_by(Sample) %>%
    summarize(nMale = sum(CellPredictedSex=="Male"), 
    	nFemale = sum(CellPredictedSex=="Female"), 
    	nCells = n()
    ) %>%
    ungroup() %>%
    mutate(PredictedSex = case_when(
    	nMale / nCells > 0.8 ~ "Male",
    	nFemale / nCells > 0.8 ~ "Female",
    	T ~ "other"
    	)
    ) %>%
    inner_join(knownSex,by="Sample") %>%
    summarize(nEqual = sum(PredictedSex==Sex), nNotEqual = sum(PredictedSex != Sex))

```

## Since everything matches we now add Sex as a metadata label

```{r}
knownSexByCell <- enframe(merged.18279.skin.singlets$Sample,name="bc",value="Sample") %>% 
	inner_join(knownSex,by="Sample") %>% 
	pull(Sex)
	
merged.18279.skin.singlets <- AddMetaData(merged.18279.skin.singlets,knownSexByCell,col.name="Sex")

```

## Plot UMAP of sex

```{r}
DimPlot(merged.18279.skin.singlets,reduction = "umap.harmony", group.by = "Sex")
```

## Save the updated object

```{r}
saveRDS(merged.18279.skin.singlets,file="Skin_scRNA_Part3.rds")
```

## Get session info

```{r}
sessionInfo()
```
