# bulkTCRseq: IMP-expanded clones Repertoire Fraction {sec-pbmc-imp-expanded-ridgeplot-barplot}

## Set up workspace
```{r}
# Libraries
library(dplyr)
library(tidyverse)
library(Seurat)
library(ggvenn)
library(ggridges)
```

## Load patient colors
```{r}
pt_cols <- readRDS("Part0_Patient_Color.rds")
pt_fill <- readRDS("Part0_Patient_Fill.rds")
```

## Open bulkTCR typed clones
```{r}
p101_betas_typed <- read.csv("p101_betas_merged_typed_Part3.csv")
p103_betas_typed <- read.csv("p103_betas_merged_typed_Part3.csv")
p104_betas_typed <- read.csv("p104_betas_merged_typed_Part3.csv")
p105_betas_typed <- read.csv("p105_betas_merged_typed_Part3.csv")
p106_betas_typed <- read.csv("p106_betas_merged_typed_Part3.csv")
p108_betas_typed <- read.csv("p108_betas_merged_typed_Part3.csv")
p109_betas_typed <- read.csv("p109_betas_merged_typed_Part3.csv")
p110_betas_typed <- read.csv("p110_betas_merged_typed_Part3.csv")
p111_betas_typed <- read.csv("p111_betas_merged_typed_Part3.csv")

betas_typed <- list(p101_betas_typed = p101_betas_typed,
                    p103_betas_typed = p103_betas_typed,
                    p104_betas_typed = p104_betas_typed,
                    p105_betas_typed = p105_betas_typed,
                    p106_betas_typed = p106_betas_typed,
                    p108_betas_typed = p108_betas_typed,
                    p109_betas_typed = p109_betas_typed,
                    p110_betas_typed = p110_betas_typed,
                    p111_betas_typed = p111_betas_typed)
```

## Compare the distribution of IMP-expanded vs non-expanded clones
```{r}
# Collect the columns of interest and rename
select_postvax_cols <- function(df){
  postvax <- df[,grep(".*postvax($|_umi)", colnames(df))]
  colnames(postvax) <- c("postvax_freq", "postvax_umi")
  df <- df[, c("Beta_clonotype", "Type", "Patient", "expanded_local_min")]
  df <- cbind(df, postvax)
  return(df)
}

postvax_tp.list <- lapply(betas_typed, select_postvax_cols)

postvax_tp <- do.call(rbind, postvax_tp.list)
```

## Plot the frequency of IMP-expanded clones after vaccination for Fig 3H middle
```{r, fig.width = 4, fig.height = 5}
p1 <- postvax_tp %>%
  filter(expanded_local_min == TRUE) %>% 
  mutate(Patient = factor(Patient, levels = c("P101", "P104", "P109", "P108", "P105", "P111", "P106", "P110", "P103"))) %>%
  ggplot(aes(x = postvax_freq, y = Patient, fill = Patient)) +
  geom_density_ridges(alpha = 0.5, scale = 1) +
  scale_x_log10() +
  pt_fill +
  theme_classic() +
  xlab("Frequency of IMP-expanded \nclones after vaccination\n")

n_expanded_postvax <- postvax_tp %>%
  filter(expanded_local_min == TRUE) %>%
  group_by(Patient) %>%
  dplyr::count() %>%
  mutate(Patient = factor(Patient, levels = c("P101", "P104", "P109", "P108", "P105", "P111", "P106", "P110", "P103")),
         x = 0.08,
         n = paste0("n=",n))

p1 + geom_text(data = n_expanded_postvax, 
              aes(y = as.numeric(Patient) + 0.1, x = x, label = n),
              position=position_nudge(y= .25), 
              colour="black", 
              size=3.5)

```

## Plot clonal dominance of IMP-expanded clones after vaccination for Fig 3H right
```{r}
fct_expanded <- postvax_tp %>%
  group_by(Patient, expanded_local_min, Type) %>%
  summarize(sum = sum(postvax_freq)) %>%
  pivot_wider(names_from = "expanded_local_min", values_from = sum) %>%
  dplyr::rename("Nonexpanded" = "FALSE",
                "Expanded" = "TRUE") %>%
  replace(is.na(.), 0) %>%
  mutate(fct_expanded = Expanded/(Expanded+Nonexpanded))

fct_expanded %>%
  group_by(Type) %>%
  summarize(median(fct_expanded)*100)

p2 <- fct_expanded %>%
  mutate(Patient = factor(Patient, levels = c("P103", "P110" , "P106" , "P111" , "P105" , "P108", "P109" , "P104", "P101"))) %>%
  ggplot(aes(x = fct_expanded, y = Type, fill = Type)) +
  geom_col(color = "#920000FF", size = 1) +
  facet_wrap(facets = ~factor(Patient, ), ncol = 1, strip.position = "right") +
  scale_x_continuous(breaks=c(0, 0.5)) +
  scale_y_discrete(expand = c(0.5, 0)) +
  ylab("Fraction of the repertoire after vaccination") +
  xlab("Fraction of the repertoire\nafter vaccination") +
  theme_classic() +
  scale_fill_manual(values = c("#B3B3B3", "#CD950C", "#0072B2"), name = "Type of IMP-\nexpanded clone") +
  theme(axis.title=element_text(size=10),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

p2
```

## Get session info
```{r}
sessionInfo()
```

